\pagebreak

\section{Version History}

\input{version-history/definitions}
% \version{x} for start of version x section

\version{1.0.2}{Snake Gourd}{}

This patch addresses the following:
\begin{itemize}
\item Improvements to how afterscripts are handled.
\item Glob arguments now expand to absolute paths
      before being evaluated.
\end{itemize}

\version{1.0.1}{Snake Gourd}{}

This patch addresses the following:
\begin{itemize}
\item A bug affecting UNIX systems, where invalid
      permissions were assigned to executable files
      downloaded from a Web server.
\end{itemize}

\version{1.0.0}{Snake Gourd}{}

This is the first stable release of the application,
marking the moment the application was turned over
to the TU Delft Algorithmics group.

\breakingchanges
\begin{itemize}
\item Many structural changes to the codebase, and a
  new configuration structure.
\end{itemize}

\version{0.8.0}{Bitter Gourd}{pre-release}

Added functionality for after-scripts, resource limits,
and incremental scheduling.
Set up the codebase for maintainability.

\breakingchanges
\begin{itemize}
\item A new configuration with many new keys and
changes to existing keys.
\end{itemize}

\version{0.1.5}{Bottle Gourd}{pre-release}

An initial implementation of the \gourdtest\ application
with basic functionality for defining experiments using a
TOML file and running them locally and on Slurm.

\breakingchanges
\begin{itemize}
\item This is a new application without compatibility with
the \gourdtest\ prototype's configuration files.
\end{itemize}

\version{0.1.0 (prototype)}{gourd-test}{}

A prototype of an evaluator for the Pumpkin CP-SAT solver,
exploring the possibility of declarative configurations.

\clearpage
